Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2017-12-19 LRDraft.tex
update

Starting from the generic time evolution equation \eqref{ketTimeEvol2} and choosing the initial condition in the form 
$\ket{\psi_0} = e^{-\\i \eps_p t_0}\ket{\psi_p}$, we can write the time evolution of a (discrete) eigenvector of $\hnot$ as:
\be \lb{ketTimeEvol3} 
\ket{\psi_p(t)} = e^{-\ii \eps_p t}\left( \identity - \ii \intdef{t_0}{t}{t'} \op{U}_0(t-t') \Phi^{a.l,s.l}(t') \op{U}_0(t'-t)  \right)\ket{\psi_p} \; ,
\ee
where we have introduced the \emph{perturbing field operators}: 
\be
\Phi^{a.l.}(t) = e^{-\eta_a |t|} \hone{t} \;, \qq
\Phi^{s.l.}(t) = \theta(t-t_0) \hone{t}
\ee
that codify the effective of the perturbing Hamiltonian in the specific switch-on procedure. Making usage of elementary properties of $\theta$ functions allows us
to rewrite the former time integral as:
\be\lb{AopDef1}
\op A(t) = \intdef{-\infty}{\infty}{t'} \op{U}^{+}_0(t-t') \Phi^{a.l,s.l}(t') \op{U}^{-}_0(t'-t) \;,
\ee
so that the effect of the switch-on protocol are completely reabsorbed in the perturbing field. In the subsequent analysis we will omit the apex in the fields $\Phi$, 
unless explicitly needed, to keep notations as simple as possible. Thanks to this definition we can express the time evolution of the eigenvectors of $\hnot$ in the 
compact form:
\be\lb{ketBraTimeEvol1}
\ket{\psi_p(t)} = e^{-\ii \eps_p t}\left( \identity - \ii \op A(t)  \right)\ket{\psi_p} \; , \qq 
\bra{\psi_p(t)} = \bra{\psi_p} \left( \identity + \ii \op A(t)  \right) e^{\ii \eps_p t} \; .
\ee

@@

We introduce the \emph{density operator} defined as the sum of the densities of all the occupied states having energy $\epsilon_p$ lower than the \emph{chemical potential} 
(Fermi energy):
\be\label{rhodef1}
\dmnot = \sum_p^{n_{occ}} \ket{\psi_p} \bra{\psi_p}  \;. 
\ee
Clearly $\dmnot$ commutes with $\hnot$ and in the unperturbed case is a time-independent operator. When the perturbation is switched the system starts to evolve
and we indicate with $\dm(t)$ the modified density of the perturbed system. We introduce the shorthand notation:
\be\lb{rhodef2}
\dm(t)  =  \dmnot + \dm'(t)\;,
\ee
where $\dm'(t)$ is called the \emph{response density operator} and represents the time dependent modification of the ground state density induced by the perturbation.
Here we are assuming that the highest occupied energy level and the lowest unoccupied one will never intersect during the time evolution. Thanks to this condition, the 
linear response regime the expression of $\dm'(t)$ can be readily derived using equations \eqref{ketBraTimeEvol1} and reads:  
\be\lb{rhoTimeEvolutionDef3}
\dm'(t) = \ii \commutator{\dmnot}{\op A(t)} \;.
\ee
The equation of motion (Liouville equation) for $\dm'$ is obtained by performing the time derivative of the previous equation. The derivative of $\op A(t)$ can be 
computed starting from its definition and reads:
\be
\d_t\op A(t) = \op\Phi(t) -\\i \commutator{\hnot}{\op A(t)} \nn \;,
\ee
and using the Jacobi identity we have:
\be\lb{LiouvilleRhoDef1}
\ii \d_t \dm'(t) = \commutator{\op\Phi(t)}{\dmnot} + \ii\commutator{\hnot}{\commutator{\dmnot}{\op A(t)}} = 
\commutator{\op\Phi(t)}{\dmnot} + \commutator{\hnot}{\dm'(t)} \;.
\ee
It is useful to introduce the \emph{Liouvillian} superoperator:
\be
\sop{\mathcal L_0} \equiv \commutator{\hnot}{\cdot} \;,
\ee
so that the Liouville equation can be written as:
\be\lb{LiouvilleRhoDef2}
\ii \d_t \dm'(t) = \sop{\mathcal L_0} \dm'(t) + \commutator{\op\Phi(t)}{\dmnot} \;.
\ee

@@

>>>>

Starting from the generic time evolution equation \eqref{ketTimeEvol2} and choosing the initial condition in the form 
$\ket{\psi_0} = e^{-\\i \eps_p t_0}\ket{\psi_p}$, we can write the time evolution of a (discrete) eigenvector of $\hnot$ as:
\be \lb{ketTimeEvol3} 
\ket{\psi_p(t)} = e^{-\ii \eps_p t}\left( \identity - \ii \intdef{t_0}{t}{t'} \op{U}_0(t-t') \op\Phi^{a.l,s.l}(t') \op{U}_0(t'-t)  \right)\ket{\psi_p} \; ,
\ee
where we have introduced the \emph{perturbing field operators}: 
\be
\op\Phi^{a.l.}(t) = e^{-\eta_a |t|} \hone{t} \;, \qq
\op\Phi^{s.l.}(t) = \theta(t-t_0) \hone{t}
\ee
that codify the effective of the perturbing Hamiltonian in the specific switch-on procedure. Making usage of elementary properties of $\theta$ functions allows us
to rewrite the former time integral as:
\be\lb{AopDef1}
\op A(t) = \intdef{-\infty}{\infty}{t'} \op{U}^{+}_0(t-t') \op\Phi^{a.l,s.l}(t') \op{U}^{-}_0(t'-t) \;,
\ee
so that the effect of the switch-on protocol are completely reabsorbed in the perturbing field. In the subsequent analysis we will omit the apex in the fields $\Phi$, 
unless explicitly needed, to keep notations as simple as possible. Thanks to this definition we can express the time evolution of the eigenvectors of $\hnot$ in the 
compact form:
\be\lb{ketBraTimeEvol1}
\ket{\psi_p(t)} = e^{-\ii \eps_p t}\left( \identity - \ii \op A(t)  \right)\ket{\psi_p} \; , \qq 
\bra{\psi_p(t)} = \bra{\psi_p} \left( \identity + \ii \op A(t)  \right) e^{\ii \eps_p t} \; .
\ee

@@

We introduce the \emph{density operator} defined as the sum of the densities of all the occupied states having energy $\epsilon_p$ lower than the \emph{chemical potential} 
(Fermi energy):
\be\label{rhodef1}
\dmnot = \sum_p^{n_{occ}} \ket{\psi_p} \bra{\psi_p}  \;. 
\ee
Clearly $\dmnot$ commutes with $\hnot$ and in the unperturbed case is a time-independent operator. When the perturbation is switched the system starts to evolve
and we indicate with $\dm(t)$ the modified density of the perturbed system. We introduce the shorthand notation:
\be\lb{rhodef2}
\dm(t)  =  \dmnot + \dm'(t)\;,
\ee
where $\dm'(t)$ is called the \emph{response density operator} and represents the time dependent modification of the ground state density induced by the perturbation.
Here we are assuming that the highest occupied energy level and the lowest unoccupied one will never intersect during the time evolution. Thanks to this condition, the 
linear response regime the expression of $\dm'(t)$ can be readily derived using equations \eqref{ketBraTimeEvol1} and reads:  
\be\lb{rhoTimeEvolutionDef3}
\dm'(t) = \ii \commutator{\dmnot}{\op A(t)} \;.
\ee
The equation of motion (Liouville equation) for $\dm'$ is obtained by performing the time derivative of the previous equation. The derivative of $\op A(t)$ can be 
computed starting from its definition and reads:
\be
\d_t\op A(t) = \op\Phi(t) -\\i \commutator{\hnot}{\op A(t)} \nn \;,
\ee
and using the Jacobi identity we have:
\be\lb{LiouvilleRhoDef1}
\ii \d_t \dm'(t) = \commutator{\op\Phi(t)}{\dmnot} + \ii\commutator{\hnot}{\commutator{\dmnot}{\op A(t)}} = 
\commutator{\op\Phi(t)}{\dmnot} + \commutator{\hnot}{\dm'(t)} \;.
\ee
It is useful to introduce the \emph{Liouvillian} superoperator:
\be\lb{L0sopDef1}
\sop{\mathcal L_0} \equiv \commutator{\hnot}{\cdot} \;,
\ee
so that the Liouville equation can be written as:
\be\lb{LiouvilleRhoDef2}
\ii \d_t \dm'(t) = \sop{\mathcal L_0} \dm'(t) + \commutator{\op\Phi(t)}{\dmnot} \;.
\ee

@@

We reformulate the Liouville equation \eqref{LiouvilleRhoSelfDef2} in the frequency domain. In order to achieve this task it is
preferable to directly compute the action of the Liouville operator on the Fourier transform of the response density, since in this
way some subtleties that concerns the analytic structure in the $\omega$ complex plane can be properly taken into account.
We first define the response density in the frequency domain that, staring from \eqref{rhoTimeEvolutionDef3}, is given by:
\be\label{rhoPrimeOmegaDef1}
\dm'(\omega) \equiv \intover{t} e^{\ii \omega t} \dm'(t)
= \ii \commutator{\dmnot}{\op A(\omega)} \;,
\ee
where $\op A(\omega)$ is The Fourier transform of \eqref{AopDef1} and reads:
\be \label{AomegaDef1}
\op A(\omega) = \frac{1}{2\pi} \int_{-\infty}^\infty  \dd \omega'
\op G_0^+(\omega') \op \Phi(\omega) \op G_0^-(\omega'-\omega)\;.
\ee
The action of $\sop{\mathcal L_0}$ on $\dm'(\omega)$ follows from \eqref{L0sopDef1}:
\be\label{LiouvillianF'def1}
\sop{\mathcal L_0} \dm'(\omega) = \ii \commutator{\hnot}{\commutator{\dmnot}{\op A(\omega)}} = 
\ii \commutator{\dmnot}{\commutator{\hnot}{\op A(\omega)}}
\;,
\ee
\be\label{eoma}
\commutator{\hnot }{\op A(\omega)} = \bar{\bar\omega} \op A(\omega) - \ii \op \Phi(\omega)\;.
\ee
where $\bar{\bar\omega} = \omega + 2\ii\eta$
